---
title: "Overall correlations summarized"
author: "Gavin Douglas"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
  html_document:
    code_folding: hide
    theme: cerulean
    toc: true
    toc_float: true
---

```{r setup}
suppressPackageStartupMessages(library(circlize))
suppressPackageStartupMessages(library(cowplot))
suppressPackageStartupMessages(library(ggbeeswarm))
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(kableExtra))
suppressPackageStartupMessages(library(knitr))
suppressPackageStartupMessages(library(reshape2))
suppressPackageStartupMessages(library(grid))
```

```{r definitions}
approaches <- c("simple", "hyperG", "propr")

env_vars <- c("depth", "latitude", "longitude", "temperature", "oxygen", "salinity")

id_map = list()

for (env_var in env_vars) {
  id_map[[env_var]] <- paste(tools::toTitleCase(env_var), "(diff.)")
}

id_map[["tip_dist"]] <- "Tip dist."

id_map[["95_gene_count"]] <- "HGT (gene count)"
id_map[["99_gene_count"]] <- "HGT (gene count)"
id_map[["both_gene_count"]] <- "HGT (gene count)"
id_map[["ranger_hgt_tallies"]] <- "HGT (gene count)"

id_map[["cooccur_simple_cooccur"]] <- "Co-occur (Simple)"
id_map[["cooccur_ratio"]] <- "Co-occur (HyperG)"
id_map[["cooccur_asso"]] <- "Co-occur (propr)"

clean_ids <- function(in_vec) { 
  sapply(in_vec, function(x) { id_map[[x]] })
}

comparison_type_cols <- c("#33a02c", "#1f78b4", "#e31a1c")


prep_pairwise_tab <- function(infile) {
  in_pairwise <- read.table(infile, header = TRUE, sep = "\t", stringsAsFactors = FALSE)
  X_tallies <- table(in_pairwise$X)
  Y_tallies <- table(in_pairwise$Y)
  in_pairwise$X <- factor(in_pairwise$X, levels = names(X_tallies)[order(X_tallies, decreasing = FALSE)])
  in_pairwise$Y <- factor(in_pairwise$Y, levels = names(Y_tallies)[order(Y_tallies, decreasing = FALSE)])
  in_pairwise_mean <- aggregate(x = r ~ X + Y, FUN = mean, data = in_pairwise)
  in_pairwise_mean_wide <- reshape2::dcast(data = in_pairwise_mean, formula = X ~ Y, value.var = 'r')
  rownames(in_pairwise_mean_wide) <- in_pairwise_mean_wide[, 1]
  
  # Also get comparisons not significantly different from 0, so that these can be coloured as grey.
  unique_combos <- in_pairwise[, c("X", "Y")]
  unique_combos <- unique_combos[-which(duplicated(unique_combos)), ]
  nonsig_comparisons <- list()
  nonsig_count <- 1
  for (i in 1:nrow(unique_combos)) {
    x_subset <- unique_combos[i, "X"]
    y_subset <- unique_combos[i, "Y"]
    combo_subset <- in_pairwise[which(in_pairwise$X == x_subset & in_pairwise$Y == y_subset), ]
    wilox_combo_out <- wilcox.test(combo_subset$r)
    if (wilox_combo_out$p.value >= 0.05) {
      nonsig_comparisons[[nonsig_count]] <- c(y_subset, x_subset)
      nonsig_count <- nonsig_count + 1
    }
  }
  
  return(list(tab=in_pairwise_mean_wide[, -1],
              nonsig_comparisons=nonsig_comparisons))
}
```

# 95%

## Pairwise Spearman correlations {.tabset}

```{r read_pairwise_files_95}
mean_pairwise = list()
mean_pairwise[["BLAST"]] <- list()

for (approach in approaches) {
  infile_blast_pairwise <- paste("/mfs/gdouglas/projects/ocean_mags/coverm/network_working_allsamples/partial_corr_out_95/", approach, ".pairwise.tsv", sep = "")
  mean_pairwise[["BLAST"]][[approach]] <- prep_pairwise_tab(infile_blast_pairwise)
}

identity_sanity_checks <- c(identical(mean_pairwise[["BLAST"]]$hyperG$tab[-8, ],  mean_pairwise[["BLAST"]]$propr$tab[-8, ]),
                            identical(mean_pairwise[["BLAST"]]$simple$tab[-8, ],  mean_pairwise[["BLAST"]]$propr$tab[-8, ]))

if (length(which(! identity_sanity_checks)) > 0) { stop("Non-co-occurrence mismatches" )}

# Then realized afterwards that I should just combine the co-occurrence types into a single heatmap.
mean_pairwise_combined <- list()

BLAST_hyperG_row <- mean_pairwise[["BLAST"]]$hyperG$tab["cooccur_ratio", , drop = FALSE]
BLAST_propr_row <- mean_pairwise[["BLAST"]]$propr$tab["cooccur_asso", , drop = FALSE]

mean_pairwise_combined[["BLAST"]] <- rbind(mean_pairwise[["BLAST"]]$simple$tab, BLAST_hyperG_row)
mean_pairwise_combined[["BLAST"]] <- rbind(mean_pairwise_combined[["BLAST"]], BLAST_propr_row)
```

```{r blast_pairwise_spearman_95, fig.height=7, fig.width=8, results='asis', echo=FALSE}
for (hgt_type in names(mean_pairwise)) {
  cat("\n\n")
  
  cat("### ", hgt_type, " results", "\n\n")
  
    in_mat <- as.matrix(mean_pairwise_combined[[hgt_type]])

    in_mat_char <- in_mat
    for (i in 1:ncol(in_mat)) {
     in_mat_char[, i] <- sprintf("%.2f", in_mat[, i])
    }
    in_mat_char[in_mat_char == "NA"] <- ""
    col_fun = colorRamp2(c(-1, 0, 1), c("#6699CC", "white", "#CC3333"))

    heatmap_out <- ComplexHeatmap::Heatmap(matrix = in_mat,
                                           col = col_fun,
                                           cluster_rows = FALSE,
                                           cluster_columns = FALSE,
                                           row_split = c(" ", " ", " ", " ", " ", "  ", "   ", "     ", "     ", "     "),
                                           row_labels = clean_ids(rownames(in_mat)),
                                           column_labels = clean_ids(colnames(in_mat)),
                                           na_col = "grey50",
                                           cell_fun = function(j, i, x, y, width, height, fill) {
                                             grid.text(in_mat_char[i, j], x, y, gp = gpar(fontsize = 10))
                                           },
                                           name = 'Spearman correlation')
    
    print(heatmap_out)
    
    cat("\n\n")

}

```

## Partial Spearman correlations {.tabset}

```{r read_allsamples_tip_partial_95}
allsamples_partial_folder <- "/mfs/gdouglas/projects/ocean_mags/coverm/network_working_allsamples/partial_corr_out_95"

allsamples_simple_partial_in <- read.table(paste(allsamples_partial_folder, "simple.partial.tsv", sep = "/"), header = TRUE, sep = "\t", stringsAsFactors = FALSE)
allsamples_simple_partial_tiponly_in <- read.table(paste(allsamples_partial_folder, "simple.tiponly.partial.tsv", sep = "/"), header = TRUE, sep = "\t", stringsAsFactors = FALSE)
allsamples_simple_partial_permuted_in <- read.table(paste(allsamples_partial_folder, "simple.permuted.partial.tsv", sep = "/"), header = TRUE, sep = "\t", stringsAsFactors = FALSE)

allsamples_simple_spearman_in <- read.table(paste(allsamples_partial_folder, "simple.pairwise.tsv", sep = "/"), header = TRUE, sep = "\t", stringsAsFactors = FALSE)
allsamples_simple_spearman_in <- allsamples_simple_spearman_in[grep("cooccur", allsamples_simple_spearman_in$X), ]
allsamples_simple_spearman_in <- allsamples_simple_spearman_in[which(allsamples_simple_spearman_in$Y == "95_gene_count"), ]
allsamples_simple_spearman_in$p.val <- allsamples_simple_spearman_in$p.unc
allsamples_simple_spearman_in <- allsamples_simple_spearman_in[, colnames(allsamples_simple_partial_tiponly_in)]

allsamples_simple_partial_in$type <- "All var. controlled"
allsamples_simple_partial_tiponly_in$type <- "Tip dist. controlled"
allsamples_simple_partial_permuted_in$type <- "All var. controlled (permuted)"
allsamples_simple_spearman_in$type <- "No control"

allsamples_simple_combined <- do.call(rbind, list(allsamples_simple_partial_in, allsamples_simple_partial_tiponly_in, 
                                                  allsamples_simple_partial_permuted_in, allsamples_simple_spearman_in))
allsamples_simple_combined$Approach <- "Simple"

allsamples_hyperG_partial_in <- read.table(paste(allsamples_partial_folder, "hyperG.partial.tsv", sep = "/"), header = TRUE, sep = "\t", stringsAsFactors = FALSE)
allsamples_hyperG_partial_tiponly_in <- read.table(paste(allsamples_partial_folder, "hyperG.tiponly.partial.tsv", sep = "/"), header = TRUE, sep = "\t", stringsAsFactors = FALSE)
allsamples_hyperG_partial_permuted_in <- read.table(paste(allsamples_partial_folder, "hyperG.permuted.partial.tsv", sep = "/"), header = TRUE, sep = "\t", stringsAsFactors = FALSE)

allsamples_hyperG_spearman_in <- read.table(paste(allsamples_partial_folder, "hyperG.pairwise.tsv", sep = "/"), header = TRUE, sep = "\t", stringsAsFactors = FALSE)
allsamples_hyperG_spearman_in <- allsamples_hyperG_spearman_in[grep("cooccur", allsamples_hyperG_spearman_in$X), ]
allsamples_hyperG_spearman_in <- allsamples_hyperG_spearman_in[which(allsamples_hyperG_spearman_in$Y == "95_gene_count"), ]
allsamples_hyperG_spearman_in$p.val <- allsamples_hyperG_spearman_in$p.unc
allsamples_hyperG_spearman_in <- allsamples_hyperG_spearman_in[, colnames(allsamples_hyperG_partial_tiponly_in)]

allsamples_hyperG_partial_in$type <- "All var. controlled"
allsamples_hyperG_partial_tiponly_in$type <- "Tip dist. controlled"
allsamples_hyperG_partial_permuted_in$type <- "All var. controlled (permuted)"
allsamples_hyperG_spearman_in$type <- "No control"

allsamples_hyperG_combined <- do.call(rbind, list(allsamples_hyperG_partial_in, allsamples_hyperG_partial_tiponly_in, 
                                                  allsamples_hyperG_partial_permuted_in, allsamples_hyperG_spearman_in))
allsamples_hyperG_combined$Approach <- "HyperG"

allsamples_propr_partial_in <- read.table(paste(allsamples_partial_folder, "propr.partial.tsv", sep = "/"), header = TRUE, sep = "\t", stringsAsFactors = FALSE)
allsamples_propr_partial_tiponly_in <- read.table(paste(allsamples_partial_folder, "propr.tiponly.partial.tsv", sep = "/"), header = TRUE, sep = "\t", stringsAsFactors = FALSE)
allsamples_propr_partial_permuted_in <- read.table(paste(allsamples_partial_folder, "propr.permuted.partial.tsv", sep = "/"), header = TRUE, sep = "\t", stringsAsFactors = FALSE)

allsamples_propr_spearman_in <- read.table(paste(allsamples_partial_folder, "propr.pairwise.tsv", sep = "/"), header = TRUE, sep = "\t", stringsAsFactors = FALSE)
allsamples_propr_spearman_in <- allsamples_propr_spearman_in[grep("cooccur", allsamples_propr_spearman_in$X), ]
allsamples_propr_spearman_in <- allsamples_propr_spearman_in[which(allsamples_propr_spearman_in$Y == "95_gene_count"), ]
allsamples_propr_spearman_in$p.val <- allsamples_propr_spearman_in$p.unc
allsamples_propr_spearman_in <- allsamples_propr_spearman_in[, colnames(allsamples_propr_partial_tiponly_in)]

allsamples_propr_partial_in$type <- "All var. controlled"
allsamples_propr_partial_tiponly_in$type <- "Tip dist. controlled"
allsamples_propr_partial_permuted_in$type <- "All var. controlled (permuted)"
allsamples_propr_spearman_in$type <- "No control"

allsamples_propr_combined <- do.call(rbind, list(allsamples_propr_partial_in, allsamples_propr_partial_tiponly_in, 
                                                  allsamples_propr_partial_permuted_in, allsamples_propr_spearman_in))
allsamples_propr_combined$Approach <- "propr"

allsamples_partial_combined <- do.call(rbind, list(allsamples_simple_combined, allsamples_hyperG_combined, allsamples_propr_combined))
allsamples_partial_combined$Approach <- factor(allsamples_partial_combined$Approach, levels = c("Simple", "HyperG", "propr"))
allsamples_partial_combined$type <- factor(allsamples_partial_combined$type, levels = c("No control", "Tip dist. controlled", "All var. controlled", "All var. controlled (permuted)"))
```

### BLAST (genus and above)

```{r allsamples_ocean_spearman_partial_blast_95, fig.height=4, fig.width=6}
allsamples_partial_combined <- allsamples_partial_combined[which(! is.na(allsamples_partial_combined$r)), ]

print(ggplot(data = allsamples_partial_combined, aes(x = Approach, y = r, fill = type)) +
                              geom_quasirandom(dodge.width=.75, alpha = 0.5, col="grey80") +
                              geom_boxplot(outlier.shape = NA, alpha=0.7) +
                              theme_bw() +
                              theme(panel.grid.major.x = element_blank(),
                                    panel.grid.minor.x = element_blank(),
                                    plot.title = element_text(hjust = 0.5)) +
                              ylab("Partial Spearman's correlation coefficient") +
                              xlab("Co-occurrence approach") +
                              ggtitle("All ocean samples (BLAST - genus and above)") +
                              labs(fill="Spearman approach") +
                              scale_fill_manual(values = c("#c7733b", "#8d75ca", "#78a450", "#c16786"))
)
```

## Partial Spearman correlations (HyperG only) {.tabset}

### BLAST (genus and above)

```{r allsamples_ocean_spearman_partial_blast_hyperg_95, fig.height=5, fig.width=5}

allsamples_partial_hyperg <- allsamples_partial_combined[which(allsamples_partial_combined$Approach == 'HyperG'), ]

allsamples_partial_hyperg$type <- as.character(allsamples_partial_hyperg$type)
allsamples_partial_hyperg$type[which(allsamples_partial_hyperg$type == 'Tip dist. controlled')] <- 'Tip dist.\ncontrolled'
allsamples_partial_hyperg$type[which(allsamples_partial_hyperg$type == 'All var. controlled')] <- 'All var.\ncontrolled'
allsamples_partial_hyperg$type[which(allsamples_partial_hyperg$type == 'All var. controlled (permuted)')] <- 'All var.\ncontrolled\n(permuted)'
allsamples_partial_hyperg$type <- factor(allsamples_partial_hyperg$type,
                                         levels = c("No control", "Tip dist.\ncontrolled", "All var.\ncontrolled", "All var.\ncontrolled\n(permuted)"))

print(ggplot(data = allsamples_partial_hyperg, aes(x = type, y = r, fill = type)) +
                              geom_quasirandom(dodge.width=.75, alpha = 0.5, col="grey80") +
                              geom_boxplot(outlier.shape = NA, alpha=0.7) +
                              theme_bw() +
                              theme(panel.grid.major.x = element_blank(),
                                    panel.grid.minor.x = element_blank(),
                                    plot.title = element_text(hjust = 0.5),
                                    legend.position = "none") +
                              ylab("Correlation coefficient") +
                              xlab("Spearman approach") +
                              ggtitle("All ocean samples (BLAST - genus and above)") +
                              scale_fill_manual(values = c("#c7733b", "#8d75ca", "#78a450", "#c16786"))
)
```

# 99%

## Pairwise Spearman correlations {.tabset}

```{r read_pairwise_files_99}
mean_pairwise = list()
mean_pairwise[["BLAST"]] <- list()

for (approach in approaches) {
  infile_blast_pairwise <- paste("/mfs/gdouglas/projects/ocean_mags/coverm/network_working_allsamples/partial_corr_out_99/", approach, ".pairwise.tsv", sep = "")
  mean_pairwise[["BLAST"]][[approach]] <- prep_pairwise_tab(infile_blast_pairwise)
}

identity_sanity_checks <- c(identical(mean_pairwise[["BLAST"]]$hyperG$tab[-8, ],  mean_pairwise[["BLAST"]]$propr$tab[-8, ]),
                            identical(mean_pairwise[["BLAST"]]$simple$tab[-8, ],  mean_pairwise[["BLAST"]]$propr$tab[-8, ]))

if (length(which(! identity_sanity_checks)) > 0) { stop("Non-co-occurrence mismatches" )}

# Then realized afterwards that I should just combine the co-occurrence types into a single heatmap.
mean_pairwise_combined <- list()

BLAST_hyperG_row <- mean_pairwise[["BLAST"]]$hyperG$tab["cooccur_ratio", , drop = FALSE]
BLAST_propr_row <- mean_pairwise[["BLAST"]]$propr$tab["cooccur_asso", , drop = FALSE]

mean_pairwise_combined[["BLAST"]] <- rbind(mean_pairwise[["BLAST"]]$simple$tab, BLAST_hyperG_row)
mean_pairwise_combined[["BLAST"]] <- rbind(mean_pairwise_combined[["BLAST"]], BLAST_propr_row)
```

```{r blast_pairwise_spearman_99, fig.height=7, fig.width=8, results='asis', echo=FALSE}
for (hgt_type in names(mean_pairwise)) {
  cat("\n\n")
  
  cat("### ", hgt_type, " results", "\n\n")
  
    in_mat <- as.matrix(mean_pairwise_combined[[hgt_type]])

    in_mat_char <- in_mat
    for (i in 1:ncol(in_mat)) {
     in_mat_char[, i] <- sprintf("%.2f", in_mat[, i])
    }
    in_mat_char[in_mat_char == "NA"] <- ""
    col_fun = colorRamp2(c(-1, 0, 1), c("#6699CC", "white", "#CC3333"))

    heatmap_out <- ComplexHeatmap::Heatmap(matrix = in_mat,
                                           col = col_fun,
                                           cluster_rows = FALSE,
                                           cluster_columns = FALSE,
                                           row_split = c(" ", " ", " ", " ", " ", "  ", "   ", "     ", "     ", "     "),
                                           row_labels = clean_ids(rownames(in_mat)),
                                           column_labels = clean_ids(colnames(in_mat)),
                                           na_col = "grey50",
                                           cell_fun = function(j, i, x, y, width, height, fill) {
                                             grid.text(in_mat_char[i, j], x, y, gp = gpar(fontsize = 10))
                                           },
                                           name = 'Spearman correlation')
    
    print(heatmap_out)
    
    cat("\n\n")

}

```

## Partial Spearman correlations {.tabset}

```{r read_allsamples_tip_partial_99}
allsamples_partial_folder <- "/mfs/gdouglas/projects/ocean_mags/coverm/network_working_allsamples/partial_corr_out_99"

allsamples_simple_partial_in <- read.table(paste(allsamples_partial_folder, "simple.partial.tsv", sep = "/"), header = TRUE, sep = "\t", stringsAsFactors = FALSE)
allsamples_simple_partial_tiponly_in <- read.table(paste(allsamples_partial_folder, "simple.tiponly.partial.tsv", sep = "/"), header = TRUE, sep = "\t", stringsAsFactors = FALSE)
allsamples_simple_partial_permuted_in <- read.table(paste(allsamples_partial_folder, "simple.permuted.partial.tsv", sep = "/"), header = TRUE, sep = "\t", stringsAsFactors = FALSE)

allsamples_simple_spearman_in <- read.table(paste(allsamples_partial_folder, "simple.pairwise.tsv", sep = "/"), header = TRUE, sep = "\t", stringsAsFactors = FALSE)
allsamples_simple_spearman_in <- allsamples_simple_spearman_in[grep("cooccur", allsamples_simple_spearman_in$X), ]
allsamples_simple_spearman_in <- allsamples_simple_spearman_in[which(allsamples_simple_spearman_in$Y == "99_gene_count"), ]
allsamples_simple_spearman_in$p.val <- allsamples_simple_spearman_in$p.unc
allsamples_simple_spearman_in <- allsamples_simple_spearman_in[, colnames(allsamples_simple_partial_tiponly_in)]

allsamples_simple_partial_in$type <- "All var. controlled"
allsamples_simple_partial_tiponly_in$type <- "Tip dist. controlled"
allsamples_simple_partial_permuted_in$type <- "All var. controlled (permuted)"
allsamples_simple_spearman_in$type <- "No control"

allsamples_simple_combined <- do.call(rbind, list(allsamples_simple_partial_in, allsamples_simple_partial_tiponly_in, 
                                                  allsamples_simple_partial_permuted_in, allsamples_simple_spearman_in))
allsamples_simple_combined$Approach <- "Simple"

allsamples_hyperG_partial_in <- read.table(paste(allsamples_partial_folder, "hyperG.partial.tsv", sep = "/"), header = TRUE, sep = "\t", stringsAsFactors = FALSE)
allsamples_hyperG_partial_tiponly_in <- read.table(paste(allsamples_partial_folder, "hyperG.tiponly.partial.tsv", sep = "/"), header = TRUE, sep = "\t", stringsAsFactors = FALSE)
allsamples_hyperG_partial_permuted_in <- read.table(paste(allsamples_partial_folder, "hyperG.permuted.partial.tsv", sep = "/"), header = TRUE, sep = "\t", stringsAsFactors = FALSE)

allsamples_hyperG_spearman_in <- read.table(paste(allsamples_partial_folder, "hyperG.pairwise.tsv", sep = "/"), header = TRUE, sep = "\t", stringsAsFactors = FALSE)
allsamples_hyperG_spearman_in <- allsamples_hyperG_spearman_in[grep("cooccur", allsamples_hyperG_spearman_in$X), ]
allsamples_hyperG_spearman_in <- allsamples_hyperG_spearman_in[which(allsamples_hyperG_spearman_in$Y == "99_gene_count"), ]
allsamples_hyperG_spearman_in$p.val <- allsamples_hyperG_spearman_in$p.unc
allsamples_hyperG_spearman_in <- allsamples_hyperG_spearman_in[, colnames(allsamples_hyperG_partial_tiponly_in)]

allsamples_hyperG_partial_in$type <- "All var. controlled"
allsamples_hyperG_partial_tiponly_in$type <- "Tip dist. controlled"
allsamples_hyperG_partial_permuted_in$type <- "All var. controlled (permuted)"
allsamples_hyperG_spearman_in$type <- "No control"

allsamples_hyperG_combined <- do.call(rbind, list(allsamples_hyperG_partial_in, allsamples_hyperG_partial_tiponly_in, 
                                                  allsamples_hyperG_partial_permuted_in, allsamples_hyperG_spearman_in))
allsamples_hyperG_combined$Approach <- "HyperG"

allsamples_propr_partial_in <- read.table(paste(allsamples_partial_folder, "propr.partial.tsv", sep = "/"), header = TRUE, sep = "\t", stringsAsFactors = FALSE)
allsamples_propr_partial_tiponly_in <- read.table(paste(allsamples_partial_folder, "propr.tiponly.partial.tsv", sep = "/"), header = TRUE, sep = "\t", stringsAsFactors = FALSE)
allsamples_propr_partial_permuted_in <- read.table(paste(allsamples_partial_folder, "propr.permuted.partial.tsv", sep = "/"), header = TRUE, sep = "\t", stringsAsFactors = FALSE)

allsamples_propr_spearman_in <- read.table(paste(allsamples_partial_folder, "propr.pairwise.tsv", sep = "/"), header = TRUE, sep = "\t", stringsAsFactors = FALSE)
allsamples_propr_spearman_in <- allsamples_propr_spearman_in[grep("cooccur", allsamples_propr_spearman_in$X), ]
allsamples_propr_spearman_in <- allsamples_propr_spearman_in[which(allsamples_propr_spearman_in$Y == "99_gene_count"), ]
allsamples_propr_spearman_in$p.val <- allsamples_propr_spearman_in$p.unc
allsamples_propr_spearman_in <- allsamples_propr_spearman_in[, colnames(allsamples_propr_partial_tiponly_in)]

allsamples_propr_partial_in$type <- "All var. controlled"
allsamples_propr_partial_tiponly_in$type <- "Tip dist. controlled"
allsamples_propr_partial_permuted_in$type <- "All var. controlled (permuted)"
allsamples_propr_spearman_in$type <- "No control"

allsamples_propr_combined <- do.call(rbind, list(allsamples_propr_partial_in, allsamples_propr_partial_tiponly_in, 
                                                  allsamples_propr_partial_permuted_in, allsamples_propr_spearman_in))
allsamples_propr_combined$Approach <- "propr"

allsamples_partial_combined <- do.call(rbind, list(allsamples_simple_combined, allsamples_hyperG_combined, allsamples_propr_combined))
allsamples_partial_combined$Approach <- factor(allsamples_partial_combined$Approach, levels = c("Simple", "HyperG", "propr"))
allsamples_partial_combined$type <- factor(allsamples_partial_combined$type, levels = c("No control", "Tip dist. controlled", "All var. controlled", "All var. controlled (permuted)"))
```

### BLAST (genus and above)

```{r allsamples_ocean_spearman_partial_blast_99, fig.height=4, fig.width=6}
allsamples_partial_combined <- allsamples_partial_combined[which(! is.na(allsamples_partial_combined$r)), ]

print(ggplot(data = allsamples_partial_combined, aes(x = Approach, y = r, fill = type)) +
                              geom_quasirandom(dodge.width=.75, alpha = 0.5, col="grey80") +
                              geom_boxplot(outlier.shape = NA, alpha=0.7) +
                              theme_bw() +
                              theme(panel.grid.major.x = element_blank(),
                                    panel.grid.minor.x = element_blank(),
                                    plot.title = element_text(hjust = 0.5)) +
                              ylab("Partial Spearman's correlation coefficient") +
                              xlab("Co-occurrence approach") +
                              ggtitle("All ocean samples (BLAST - genus and above)") +
                              labs(fill="Spearman approach") +
                              scale_fill_manual(values = c("#c7733b", "#8d75ca", "#78a450", "#c16786"))
)
```

## Partial Spearman correlations (HyperG only) {.tabset}

### BLAST (genus and above)

```{r allsamples_ocean_spearman_partial_blast_hyperg_99, fig.height=5, fig.width=5}

allsamples_partial_hyperg <- allsamples_partial_combined[which(allsamples_partial_combined$Approach == 'HyperG'), ]

allsamples_partial_hyperg$type <- as.character(allsamples_partial_hyperg$type)
allsamples_partial_hyperg$type[which(allsamples_partial_hyperg$type == 'Tip dist. controlled')] <- 'Tip dist.\ncontrolled'
allsamples_partial_hyperg$type[which(allsamples_partial_hyperg$type == 'All var. controlled')] <- 'All var.\ncontrolled'
allsamples_partial_hyperg$type[which(allsamples_partial_hyperg$type == 'All var. controlled (permuted)')] <- 'All var.\ncontrolled\n(permuted)'
allsamples_partial_hyperg$type <- factor(allsamples_partial_hyperg$type,
                                         levels = c("No control", "Tip dist.\ncontrolled", "All var.\ncontrolled", "All var.\ncontrolled\n(permuted)"))

print(ggplot(data = allsamples_partial_hyperg, aes(x = type, y = r, fill = type)) +
                              geom_quasirandom(dodge.width=.75, alpha = 0.5, col="grey80") +
                              geom_boxplot(outlier.shape = NA, alpha=0.7) +
                              theme_bw() +
                              theme(panel.grid.major.x = element_blank(),
                                    panel.grid.minor.x = element_blank(),
                                    plot.title = element_text(hjust = 0.5),
                                    legend.position = "none") +
                              ylab("Correlation coefficient") +
                              xlab("Spearman approach") +
                              ggtitle("All ocean samples (BLAST - genus and above)") +
                              scale_fill_manual(values = c("#c7733b", "#8d75ca", "#78a450", "#c16786"))
)
```

# Session info {.tabset}

## Hide

## Show

```{r show_info}
sessionInfo()
```
